\section{Motivation and Gap}

Memory updating in the hippocampal-entorhinal system is increasingly described as a problem of retrieval selection rather than storage replacement. Earlier traces are not eliminated when new learning occurs. Distinct engram populations can persist, compete, and become differentially accessible according to context, inhibitory tone, and network state \cite{arxiv250601659,arxiv220604924,arxiv171109876}. This view shifts the central question from how a stored value is overwritten to how one available trace is allowed to dominate behavior during retrieval.

The medial septum is a strong candidate regulator of this selection process because septal GABAergic neurons organize hippocampal and entorhinal rhythmic state. Septal output is associated with theta-promoting circuit dynamics, state transitions between active processing and replay, and long-range timing control within the hippocampal formation \cite{arxiv151001100}. Related work on hippocampal-entorhinal information-sharing networks shows that ensemble structure is not fixed. Core-periphery organization can change across temporal windows, providing a network-level substrate for switching between active computational states \cite{arxiv200106371}.

Recent search results therefore suggest a specific mechanistic gap. The medial septum-to-medial entorhinal cortex pathway is repeatedly implicated in oscillatory control, arousal-sensitive coordination, and behaviorally relevant switching. However, the intermediate transformation is not yet specified at the level needed for a retrieval model. The unresolved step is how septal GABAergic input is converted within medial entorhinal cortex into a timing signal that can reorganize downstream CA1 output. A circuit explanation requires more than the statement that septal activity supports theta. It must specify how inhibitory microcircuits in medial entorhinal cortex alter the timing of entorhinal output, and how that timing shift changes the CA1 ensemble state expressed during retrieval.

The gap addressed here is therefore narrow. The relevant mechanism is not broad memory enhancement, generalized arousal, or overall locomotor facilitation. The target mechanism is timing-based retrieval selection after sequential learning in the same context.
